---
title: "Effect of calcium addition on Sugar Maple seedling growth"
author: "Siena Blier"
date: "2025-09-11"
output:
  pdf_document:
    fig_caption: true
header-includes:
  - \usepackage{float}
bibliography: references.bib
csl: ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "H")

# Load libraries
library(car)
library(dplyr)
library(ggplot2)
```

# Introduction

I am citing a paper [@juice2006a]. And another one @patykowski2018.

# Methods

I got the data from [@juice2019].

```{r import.data}
clean.data <- read.csv("../02_outdata/hbr-maples_clean-dataset.csv")
```

We checked that the assumptions of a t-test were met.

1.  Stem dry mass and leaf dry mass are both **continuous** variables.

```{r look.continuous}
head(clean.data$stem_dry_mass)
head(clean.data$leaf_dry_mass)
```

2.  We assume that @juice2019 carried out **random sampling**.
3.  The assumption of **homoscedasticity** is not respected, so we use a t-test with **unequal variance**

```{r check.var}

# Check equal variance for stem dry mass
aggregate(stem_dry_mass~watershed, data=clean.data, var)
leveneTest(clean.data$stem_dry_mass, clean.data$watershed)

# Check equal variance for leaf dry mass
aggregate(leaf_dry_mass~watershed, data=clean.data, var)
leveneTest(clean.data$leaf_dry_mass, clean.data$watershed)
```

4.  The data are not **normally distributed**, but because we're dealing with a fairly large sample size (n = 359), we will still conduct the t-test.

```{r check.normal}

# Check normality of stem dry mass for reference samples

clean.data.ref <- clean.data %>% 
  filter(watershed == "Reference")

shapiro.test(clean.data.ref$stem_dry_mass)
shapiro.test(clean.data.ref$leaf_dry_mass)

clean.data.w1 <- clean.data %>% 
  filter(watershed == "W1")

shapiro.test(clean.data.w1$stem_dry_mass)
shapiro.test(clean.data.w1$leaf_dry_mass)

rm(clean.data.ref, clean.data.w1)

```

# Results

A first look at the distribution of the data suggested there may be differences in stem and leaf dry mass between reference and calcium-treated seedlings.

```{r histograms, fig.cap= "Distribution of stem dry mass (left) and leaf dry mass (right) of seedlings in reference (red) and calcium treated (blue) samples (n = 359).", fig.show="hold", out.width="50%"}

# Histogram of stem dry mass
ggplot(clean.data, aes(x = stem_dry_mass, fill = watershed)) +
  geom_histogram(alpha = 0.6, position = "identity", bins = 30) +
  labs(x = "Value", y = "Frequency") +
  theme_classic()

# Histogram of leaf dry mass
ggplot(clean.data, aes(x = leaf_dry_mass, fill = watershed)) +
  geom_histogram(alpha = 0.6, position = "identity", bins = 30) +
  labs(x = "Value", y = "Frequency") +
  theme_classic()
```

We found that stem and leaf dry mass differed significantly between calcium treated and reference samples.

```{r t.test}

# Test for differences in stem dry mass
t.test(stem_dry_mass ~ watershed, data = clean.data, var.equal = FALSE) # Use Welch's Two Sample t-test (unequal variances)

# Test for differenecs in leaf dry mass
t.test(leaf_dry_mass ~ watershed, data = clean.data, var.equal = FALSE) # Use Welch's Two Sample t-test (unequal variances)
```

# Discussion

A citation for the discussion @kobe2002.

# References
